---
title: "Data Analytics Using R"
author: "Nandha Kishore G (23114)"
format: html
editor: visual
---

```{r}
install.packages("ggplot2")
library(ggplot2)

```

```{r}
# Load the dataset
Student_Performance <- read.csv("D:/r & python/students data/Student_Performance.csv")
```

```{r}
Student_Performance <- read.csv("D:/r & python/students data/Student_Performance.csv")
```

```{r}
# Build the multiple linear regression model
model <- lm(Performance.Index ~ Hours.Studied + Previous.Scores + Extracurricular.Activities, data = Student_Performance)

# Print the summary of the model
summary(model)
```

```{r}
# Check assumptions and diagnostic plots
plot(model)
```

```{r}
# Check variable significance
summary(model)$coefficients
```

```{r}
# Check multicollinearity
vif <- car::vif(model)
correlation_matrix <- cor(Student_Performance[, -c(3)])
```

```{r}
# Plot residuals vs fitted values
plot(model, which = 1)
```

```{r}

```

```{r}
# Create ggplot for Hours.Studied
ggplot(Student_Performance, aes(x = Hours.Studied, y = Performance.Index)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(title = "Hours Studied vs Performance Index")
  
```

```{r}
# Create ggplot for Previous.Scores
ggplot(Student_Performance, aes(x = Previous.Scores, y = Performance.Index)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(title = "Previous Scores vs Performance Index")
```

```{r}
# Create ggplot for Extracurricular.Activities
ggplot(Student_Performance, aes(x = Extracurricular.Activities, y = Performance.Index)) +
  geom_boxplot() +
  labs(title = "Extracurricular Activities vs Performance Index")
```

```{r}
# Create ggplot for Sleep.Hours
ggplot(Student_Performance, aes(x = Sleep.Hours, y = Performance.Index)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(title = "Sleep Hours vs Performance Index")

```

```{r}
# Create ggplot for Sample.Question.Papers.Practiced
ggplot(Student_Performance, aes(x = Sample.Question.Papers.Practiced, y = Performance.Index)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(title = "Sample Question Papers Practiced vs Performance Index")
```

```{r}
# Assume you have new data for prediction
new_data <- data.frame(
  Hours.Studied = c(6, 8, 5),
  Previous.Scores = c(70, 85, 60),
  Extracurricular.Activities = c("Yes", "Yes", "No"),
  Sleep.Hours = c(7, 9, 6),
  Sample.Question.Papers.Practiced = c(3, 5, 2)
)
```

```{r}
# Make predictions
predictions <- predict(model, newdata = new_data)
```

```{r}
print(predictions)
```

```{r}
# Extract coefficient values from the model summary
coefficients <- coef(summary(model))

# Print the coefficients
print(coefficients)
```

## Model Interpretation

### Coefficients Interpretation

The coefficients of the regression model represent the change in the response variable Performance Index

-   **Hours Studied**: For each additional hour studied, the Performance Index is expected to increase by 2.85 units.
-   **Previous Scores**: For each additional point in previous scores, the Performance Index is expected to increase by 1.01 units.
-   **Extracurricular Activities (Yes)**: Compared to students not participating in extracurricular activities, students who participate are expected to have a 0.59 increase in Performance Index.

### Practical Interpretation

-   Increasing the number of hours studied is associated with improved performance, indicating the importance of study time.
-   Higher previous scores correlate with higher performance in the current exam.
-   Participation in extracurricular activities is associated with a negative impact on performance.

These interpretations provide insights into the factors influencing student performance and can guide educational strategies and interventions.

## Conclusion

### Summary of Findings

In this analysis, we explored the relationship between various predictor variables and student performance. The multiple linear regression model revealed several important insights:

-   **Hours Studied**: A positive association was found between the number of hours studied and student performance, highlighting the importance of dedicated study time.
-   **Previous Scores**: Higher previous scores were associated with better performance in the current exam, indicating the continuity of academic achievement.
-   **Extracurricular Activities**: Participation in extracurricular activities showed a positive impact on student performance, suggesting the holistic benefits of student engagement beyond academics.

### Practical Implications

The findings of this analysis have several practical implications for educational institutions, educators, and students:

-   **Educational Strategies**: Insights from the regression analysis can inform the development of effective educational strategies focused on optimizing study time, leveraging previous achievements, and promoting holistic student engagement.
-   **Test Preparation Strategies**: Implementing test preparation strategies, such as regular practice exams, can enhance student readiness and confidence for assessments.
-   **Individualized Support**: Tailoring interventions and support mechanisms to address individual student needs can foster a more inclusive and supportive learning environment.

### Closing Remarks

In conclusion, this regression analysis contributes to our understanding of the multifaceted factors influencing student performance. By translating insights into actionable strategies and initiatives, educational stakeholders can work towards creating a supportive and enriching learning environment that nurtures the academic success and well-being of all students.
